---
title: "Analysis of Crypto Currencies"
author: "Quentin Picard"
date: "October 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align = "center", include=FALSE, echo=FALSE, warning=FALSE, message=FALSE, cache=T)
```

```{r Libraries}
library(data.table)
library(dplyr)
library(ggplot2)
library(ggvis)
library(ggfortify)
library(xts)
library(quantmod)
library(PerformanceAnalytics)
library(forecast)
```


```{r Data_load, echo=F, cache=F, warning=F, eval=T}
raw.crypto <- fread("./crypto.csv", stringsAsFactors = F)
crypto <- as.data.table(raw.crypto)
crypto <- crypto[!is.na(date)] # remove 4 NA records
crypto <- na.omit(crypto)
```

# Data 
The data file studied here contains OHLC (Open-High-Low-Close) data for about 1100 crypto currencies during the last 5 years. The data was scrapped from https://coinmarketcap.com/ by [Jvent](https://www.kaggle.com/jessevent) and uploaded to Kaggle.

Source file: 
[link](https://www.kaggle.com/jessevent/all-crypto-currencies) 

Here is a sample of the data:
```{r Data_Exploration, include=T, echo=F}
crypto %>% head(5)
```

```{r}
str(crypto)
```

```{r variable_formats}
# Columns format fix
crypto[,date:=as.Date(date)]
crypto$volume <- as.numeric(crypto$volume)
crypto$market <- as.numeric(crypto$market)
```

```{r Duplicates}
# Both coin/ symbol columns have duplicates (e.g. "Bitcoin"/ "ABC")
crypto$symbol <- paste(crypto$symbol, " (", crypto$coin, ")", sep = "")
```

The file contains `r length(levels(as.factor(crypto$symbol)))` different coins. Here are the first 20:
```{r Coin_sample, include=T}
# List of coins
l <- levels(as.factor(crypto$symbol)) %>% head(20) 
for (i in l) {
  cat(i)
  cat("\n")
}
```

```{r NAs analysis}
# analysis of NAs
NA.count <- sapply(crypto, function(x) sum(is.na(x)))
NA.percentage <- sapply(crypto, function(x) 100*sum(is.na(x))/length(x))
NA.summary = cbind(NA.count[NA.count>0], NA.percentage[NA.count>0])
NA.summary[order(desc(NA.summary[,2])),] # show by decreasing % of NA's (matrix sorting)
NA.summary[order(-(NA.summary[,2])),] # - means DESC
# Conclusion: all cells have numeric values (but there are zeros!)
```

```{r MissingDates}
# holes in dates
test <- crypto %>% 
  group_by(date) %>%
  summarize() 

test$date <- c(0,diff(test$date)) # date difference between 2 rows
dim(test[test$date != 1,]) # no holes (at least one coin per date)
```

```{r MissingMarketCap}
# 0 market caps
crypto %>%
  filter(market==0) %>% 
  filter(close>0) %>% # price not 0
  filter(volume>1000) %>% # some volume
  dim
# these are probably missing data points
```
### Market Capitalisation Distribution

Looking at the distribution of coin market caps, we observe a large skew towards smaller valuations. 
```{r MarketCap_Hist, include=T, cache=T}
# "Present" Market Cap histogram (latest day)
crypto[crypto$date=="2017-09-28",] %>% 
  ggplot(aes(x = market)) + 
  geom_histogram(binwidth = 1e6) + coord_cartesian(xlim = c(0, 1e8))


# Taking log and removing coins without 0 market cap
crypto[crypto$date=="2017-09-28" & crypto$market != 0,] %>% 
  ggplot(aes(x = log(market+1))) + 
  geom_histogram(binwidth = 0.1) #+ coord_cartesian(xlim = c(55, 70))

```

```{r SmallMarketCap}
null_mc <- dim(crypto[crypto$date=="2017-09-28" & crypto$market==0,])[1]
null_mc
small_mc <- dim(crypto[crypto$date=="2017-09-28" & crypto$market<=1e6,])[1]
small_mc
# Percentages
null_mc_perc <- round(100*dim(crypto[crypto$date=="2017-09-28" & crypto$market==0,])[1]/length(levels(as.factor(crypto$symbol))), 2)
null_mc_perc
small_mc_perc <- round(100*dim(crypto[crypto$date=="2017-09-28" & crypto$market<=1e6,])[1]/length(levels(as.factor(crypto$symbol))), 2)
small_mc_perc
```
In fact `r null_mc` (`r null_mc_perc`%) of coins have no market capitalisation as of September 28th, 2017. `r small_mc` (`r small_mc_perc`%) have a market cap below $1 million.

### Coins creation
The following graph shows the constant rise of reported coins. We observe some volatility in the number of new coins in last few months. One factor was China's decision to ban initial coin offerings (ICOs) on September, 4th.
Today the pace of new ICOs globally is still strong.
```{r Montly_coin_count, include=T}
#View(crypto)
crypto_active <- crypto[, 
                        .(date, month=format(date,"%Y%m"),
                          first_month=ifelse(date==min(date),"yes", "no"), 
                          last_month=ifelse(date==max(date) & format(date, "%Y%m")!="201709","yes", "no")),
                        by=.(coin)] %>%
  ungroup() %>%
  group_by(month) %>%
  summarise(active=n_distinct(coin), 
            initiations=sum(first_month=="yes"), 
            discontinuations=sum(last_month=="yes"))

#View(crypto_active)
crypto_active %>%
  ggplot(aes(x=month, y=value)) + 
  geom_line(aes(y=active, col="Monthly active coins"), group=1) +
  geom_line(aes(y=initiations, col="Coin initiations"), group=1) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  coord_cartesian(ylim = c(0, 1100)) +
  ggtitle("Coin initiations (coinmarketcap.com)") +
  theme(plot.title = element_text(hjust = 0.5))

```
```{r}
# Consistency check (despite holes in data per next chunck)
sum(crypto_active$initiations-crypto_active$discontinuations) == crypto_active$active[dim(crypto_active)[1]]
```

```{r}
# Analysis: showing some coins are not reported on all months

# unique coins for 2013-11
coins201311 <-crypto %>% 
  mutate(month=format(date,"%Y%m")) %>%
  subset(select=c(12, 1:11)) %>% # put new month column at the beginning
  distinct(month, coin) %>%          # useful to remove extra rows !
  arrange(month, coin) %>%
  filter(month=="201311")

coins201310 <-crypto %>% 
  mutate(month=format(date,"%Y%m")) %>%
  subset(select=c(12, 1:11)) %>% # put new month column at the beginning
  distinct(month, coin) %>% # useful ! 
  arrange(month, coin) %>%
  filter(month=="201310")

difference <- coins201311 %>% anti_join(coins201310, by = "coin")

newcoins201311 <- crypto %>%
  filter(grepl("2013-11", date)) %>%
  distinct(coin)

difference %>% anti_join(newcoins201311, by = "coin") # Bottlecaps not explained by coin initiations

# showing gap months for Bottlecaps (e.g. 2013-10)
crypto %>% 
  mutate(month=format(date,"%Y%m")) %>%
  subset(select=c(12, 1:11)) %>% # put new month column at the beginning
  distinct(month, coin) %>% # useful ! 
  arrange(month, coin) %>%
  filter(coin=="Bottlecaps")
```

```{r holdings_cash, cache=T}

# setting a volume threshold at which to buy
volume_threshold <- 100000

# min with default trade_date 
Robustmin <- function(x) {if (length(x)>0) min(x) else Sys.Date()+1} # trade date when no trade was made: tomorrow

# define trade date
crypto_trade_date <- crypto %>% 
  group_by(coin) %>%  
  arrange(date, coin) %>%
  mutate(trade_date=Robustmin(date[volume>volume_threshold])) # add trade_date

# get trade price at trade date (close of the day). Check that a date exist and that the price is not 0
crypto_trade_price <- crypto_trade_date %>%
  mutate(trade_price=ifelse(length(close[date==trade_date])==1,ifelse(close[date==trade_date]!=0,close[date==trade_date],Inf),Inf))

# calculate coin holding on any given date
crypto_coin_holding <- crypto_trade_price %>%
  mutate(holding= ifelse(date>=trade_date, close / trade_price, 0)) # holding should be 0 before trade happens

# holdings per coin and total with cash
holdings_cash <- crypto_coin_holding %>%
  group_by(date) %>%
  mutate(cash=-sum(trade_price!=Inf & date >= trade_date)) %>%
  mutate(Total_holdings=sum(holding)+cash) %>%
  mutate(holding_coin_perc=holding/(Total_holdings-cash)*100) %>%  # % of coin assets only (excluding cash)
  arrange(desc(date))
```

### Coins Market Cap graph

This graph shows the evolution of each coin market cap over time. 
```{r MarketCapGraph, include=T}
# Graph - market cap evolution
options("scipen" = 10)
# TO DO: https://stackoverflow.com/questions/32583129/ggvis-multiple-lines-with-tooltips to fix date format
symbol_tooltip <- function(x) {
  if (is.null(x)) return(NULL)
  # cat(x[[1]], x[[2]], x[[3]])
  
  paste0(names(x)[3],": <b>", x[[3]], "</b><br>", # Coin
         "Market Cap: ", format(x[[2]],big.mark=",", trim=TRUE), "<br>", # Market cap
         names(x)[1], ": ", as.Date(x[[1]]/86400000, origin='1970-01-01'), "<br>")
}

# Coin filter
coin_filter <- holdings_cash %>%
  group_by(symbol) %>%
  summarise(MC_threshold=max(market)) %>% 
  filter(MC_threshold>1e9)

#  ************ Coin Market Cap liens with tooltip ************
# Uses both lines and points (transparent) so point give right date
inner_join(holdings_cash, coin_filter, by= "symbol") %>%
  ggvis(x = ~date, y = ~market, stroke = ~symbol) %>%
  layer_lines(strokeWidth := 1) %>%
  layer_points(size = 1, fill = ~symbol, opacity := 0) %>%
  # add_tooltip(symbol_tooltip, on = "hover") %>%  #commented to be able to run all file
  hide_legend('fill') %>%
  hide_legend('stroke')
# add_axis("x", title = "Date") %>%  # doesn't work? unused argument (title = "Date")
# add_axis("y", title = "Market Cap") %>%
```

Since a few coins were clearly dominant we will now represent their relative share of the overall crypto currency pie as measured by market cap.
```{r MarketDominance, include=T}
# Stacked Area Chart

# coin ordering - decreasing Market Cap (last day)
coin_order2 <- inner_join(holdings_cash, coin_filter, by= "symbol") %>% 
  filter(date=="2017-09-28") %>%
  group_by(symbol) %>%
  arrange(date, desc(market)) 

# Plot
inner_join(holdings_cash, coin_filter, by= "symbol") %>% 
  group_by(symbol) %>%
  arrange(date, desc(market)) %>%
  ungroup() %>%
  group_by(date) %>%
  mutate(market_prop = market/sum(market)) %>% 
  ggplot(aes(x=date, y=market_prop, fill=factor(symbol, levels = as.list(coin_order2)$symbol))) + 
  geom_area(alpha=1 , size=1) #+ 
  scale_fill_brewer(palette="PuBu")

```

### Return distribution

In the graph below, all coins prices were normalized to start trade at $1 on April 28th, 2013.
Plotting prices or log prices for all coins isn't a viable option here. The coins should be clustered so that it's possible to read the data. This analysis is outside the scope of this document.
```{r Graph with normalized start date & price, include=T}

crypto_initiation_date <- crypto %>%  group_by(coin) %>%  mutate(initiation_date=min(date)) # add initiation date

crypto_initiation_date_price <- crypto_initiation_date %>%
  mutate(initiation_price=close[date==initiation_date])  # condition on different line in group ! very useful !

crypto_norm <- crypto_initiation_date_price %>%
  mutate(relative_date=(date-(initiation_date-as.Date("2013-04-28"))), # calculate relative date to normalize starts
         relative_price=ifelse(initiation_price!=0, close/initiation_price, close) ) 
#View(crypto_norm)

crypto_norm_volume <- crypto_norm %>% group_by(coin) %>% mutate(avg_volume=mean(volume)) # average volume per coin
#View(crypto_norm_volume)

library(gridExtra)

p1 <- crypto_norm_volume[,] %>%
  filter(avg_volume > 100000) %>%
  ggplot() + 
  geom_line(aes(x=relative_date, y=log(relative_price+1), group=symbol)) 
p2 <- crypto_norm_volume[,] %>%
  filter(avg_volume > 100000) %>%
  ggplot() + 
  geom_line(aes(x=relative_date, y=log(relative_price+1), group=symbol)) + coord_polar()

#grid.arrange(p1, p2, ncol = 2, top = "Log coin price series")
p1

```

```{r dollar averaging}

# Tests on trading date
crypto_trade_date[crypto_trade_date$coin=="808Coin" & 
                    (crypto_trade_date$date == crypto_trade_date$trade_date |
                       crypto_trade_date$date == crypto_trade_date$trade_date -1 |
                       crypto_trade_date$date == crypto_trade_date$trade_date +1),]

# Tests on trading price
crypto_trade_price[crypto_trade_price$symbol=="FLDC" & 
                    (crypto_trade_price$date == crypto_trade_price$trade_date |
                       crypto_trade_price$date == crypto_trade_price$trade_date -1 |
                       crypto_trade_price$date == crypto_trade_price$trade_date +1),]
# Conclusion: need to look into coins with volume > threshold but price = 0 (e.g. FLDC)
```

```{r Stats}
# % of coins bought
perc <- dim(crypto_trade_date[crypto_trade_date$trade_date!=Sys.Date()+1,])[1]/dim(crypto_trade_date)[1]*100
cat("Percentage of coins bought (based on volume threshold): \n", perc, "\n")
```

```{r Portfolio Pie Chart prep}
# Total Holdings 
holdings <- crypto_coin_holding %>%
  group_by(date) %>%
  filter(date=="2017-09-28") %>%
  mutate(Total_holdings=sum(holding)) %>%
  mutate(holding_coin_perc=holding/Total_holdings*100) %>%
  filter(holding_coin_perc>0.5) %>%
  arrange(desc(holding_coin_perc))

holdings

```

### Trading Strategies
In order to better understand the behavior of these coins, I implemented a basic dollar averaging trading strategy. Starting 5 years ago when data is available, everytime the daily trading volume passes a certain threshold ($100,000 to start) for a given coin, I buy $1 of it and hold it forever. 
The composition of the portfolio on September 28th, 2017 is shown below:

```{r DollarAveragingPieChart, include=T, fig.width=12, fig.height=8,}
# coin ordering for pie chart
coin_order <- crypto_coin_holding %>%
  group_by(date) %>%
  filter(date=="2017-09-28") %>%
  mutate(Total_holdings=sum(holding)) %>%
  mutate(holding_coin_perc=holding/Total_holdings*100) %>%
  filter(holding_coin_perc>0.5) %>%
  arrange(holding_coin_perc) %>%
  ungroup() %>%
  select(coin)

# need to reorder fill factor levels in inverse order of appearance (see next chunk)
holdings %>% ggplot(aes(x = "", y= holding_coin_perc, fill=factor(coin, levels = as.list(coin_order)$coin))) + 
  geom_bar(stat = "identity") + 
  coord_polar(theta = "y") +
  ggtitle("$ Averaging Final Portfolio") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_discrete(name = "Top coins in Dollar Averaging Portfolio (smallest first)")
```

Conclusion: 50% of the portfolio is made of just 6 coins. Stratis, the biggest position, went up 200,000 times. The next step would be to check if those coins indeed could have been purchased at the initial 'low' prices reported on Coinmarketcap.com. Historical exchange data would be needed for this.

```{r PortfolioReturn}

# Return Percentage
return <- -100 *(holdings_cash$Total_holdings[1] - holdings_cash$cash[1]) / holdings_cash$cash[1]
cat("Percentage return: ", return, "\n")
```
The value of the portfolio over time (net of the cash investment represented in red) is plotted below:
```{r PortfolioValue, include=T}
holdings_cash %>%
  filter(holding_coin_perc>1) %>%
  ggplot(aes(x=date)) + 
  geom_line(aes(y=holding, group=symbol)) +
  geom_line(aes(y=Total_holdings, color = "Total Holdings (coins + cash)")) + 
  geom_line(aes(y=cash, color = "Cash invested"))
```

```{r Coin holdings ~ volume}

Holdings_from_volume <- function(crypto, volume_threshold=100000, symbols_to_exclude=list("")) {
  #returns a list with the total final holding and the percentage return on the invested cash

# define trade date
crypto_trade_date <- crypto %>% 
  group_by(coin) %>%  
  arrange(date, coin) %>%
  mutate(trade_date=Robustmin(date[volume>volume_threshold])) # add trade_date

# get trade price at trade date (close of the day). Check that a date exist and that the price is not 0
crypto_trade_price <- crypto_trade_date %>%
  mutate(trade_price=ifelse(length(close[date==trade_date])==1,ifelse(close[date==trade_date]!=0,close[date==trade_date],Inf),Inf))

# calculate coin holding on any given date
crypto_coin_holding <- crypto_trade_price %>%
  filter(!(symbol %in% symbols_to_exclude)) %>%    # add filter: symbol in arguments
  mutate(holding= ifelse(date>=trade_date, close / trade_price, 0)) # holding should be 0 before trade happens

# holdings per coin and total with cash
holdings_cash <- crypto_coin_holding %>%
  group_by(date) %>%
  mutate(cash=-sum(trade_price!=Inf & date >= trade_date)) %>%
  mutate(Total_holdings=sum(holding)+cash) %>%
  mutate(holding_coin_perc=holding/(Total_holdings-cash)*100) %>%  # % of coin assets only (excluding cash)
  arrange(desc(date))

# return total holding value (last date)
# a <- holdings_cash[holdings_cash$date==as.Date("2017-09-28"),c("Total_holdings")] # core R
total_holding <- holdings_cash %>%  distinct(date,Total_holdings)

#cat("Total holdings: \n", total_holding$Total_holdings[1], "\n")
c(as.numeric(total_holding$Total_holdings[1]),
     -100 *(holdings_cash$Total_holdings[1] - holdings_cash$cash[1]) / holdings_cash$cash[1])
}

Holdings_from_volume(crypto, 100000, "")
Holdings_from_volume(crypto, 100000, list("BTC", "ETH"))


```

Finally I plotted the value of the final portfolio as a function of the volume threshold at which I would buy each coin.
The gains get exponentially higher as the threshold diminishes. 
Same conclusion here: we would need to check with exchange data if these coins were available at such small volumes. At the very least we can conclude that this trading strategy would only be available to small players.

```{r VolumeGraph, include=T, cache=T}
# Total holdings obtained as a function of volume treshold
n <- 0:9
vol_vec <- c(1000*((3)^n))
# vol_vec
hold_vec <- sapply(vol_vec, function(x) {Holdings_from_volume(crypto, x)})
# hold_vec

df <- data.frame(vol_vec,hold_vec[1,1:length(vol_vec)],hold_vec[2,1:length(vol_vec)])
ggplot(df, aes(vol_vec)) + 
  geom_line(aes(y=hold_vec[2,1:length(vol_vec)]), colour="black") + ylab("Total Return (%)") + xlab("Buy volume threshold")
```




```{r Holdings graph}
# Commenting - Market Cap graph above is enough

#holdings_cash$id <- 1:nrow(holdings_cash) # create index so key is unique for tooltip - doesn't work

# do not use scientific notation
#options("scipen" = 10)

# Graph - holdings per coin
# holdings_cash %>%
#   group_by(coin) %>% 
#   filter(max(holding_coin_perc)>5) %>%# only coins that at some point where >5% of the total holdings
#   ggvis(x = ~date, y = ~holding, fill = ~coin) %>%
#   add_tooltip(symbol_tooltip, on = "hover") %>%
#   layer_points(size := 10, 
#                size.hover := 200,
#                fillOpacity := 1, 
#                fillOpacity.hover := 0.5) %>% 
#   hide_legend('fill') 

#  filter(date>min(holdings_cash$date) & date<max(holdings_cash$date))

```



```{r self_contained = FALSE}
# Time Series

# filter last 12 months of data
coins_12m <- holdings_cash %>%
  filter(date > (as.Date("2017-09-28")-365)) %>%
  arrange(date)

# zoo/ xts
days <- seq(min(coins_12m$date), max(coins_12m$date), by = "day")
btc <- zoo(coins_12m[coins_12m$symbol=="BTC (Bitcoin)", "close"], days)
btc.full <- zoo(coins_12m[coins_12m$symbol=="BTC (Bitcoin)", c("open", "high", "low", "close", "volume")], days) # with all info
eth <- zoo(coins_12m[coins_12m$symbol=="ETH (Ethereum)", "close"], days)

autoplot(btc, main = "Bitcoin price - Past year")
plot(btc, main = "Bitcoin price - Past year")

btc_eth <- as.xts(data.frame(btc_col = btc, 
                             eth_col = eth))

colnames(btc_eth) <- c("btc_close", "eth_close")
head(btc_eth)

plot(as.zoo(btc_eth), screens = 1, lty = 1:2, xlab = "Date", ylab = "Closing price")
legend("topleft", c("btc", "eth"), lty = 1:2, cex = 0.5)

# ts - plot works but x axis to fix
# btc_ts <- ts(coins_12m[coins_12m$symbol=="BTC", "close"],
#            start = c(2016, as.numeric(format(days[1], "%j"))),
#            frequency = 365)
# autoplot(btc_ts) # x axis format to fix

```

```{r Returns}
# calculate returns - Method #1
quantmod::Delt(btc_eth[, "btc_close"]) %>% head #  period pct. price change - is xts
quantmod::Delt(as.xts(btc[, "close"])) %>% head #  period pct. price change

# method #2 - vectorized formula
n <- nrow(btc)
apply(btc, 2, function (x) { diff(x)/x[-length(x)] }) %>% head

# For n coins (method #2)
btc_eth_returns <- as.xts(apply(btc_eth, 2, function (x) { diff(x)/x[-length(x)] }))
btc_eth_returns %>% head

plot(as.zoo(btc_eth_returns), screens = 1, lty = 1:2, xlab = "Date", ylab = "Return")
legend("topleft", c("Bitcoin", "Ether"), lty = 1:2, cex = 0.5)


```

```{r}
# boxplot BTC ~ ETH (back to vectors ? better function to use ?)
eth_decile <- cut(as.vector(btc_eth_returns[,2]), 10)
boxplot(as.vector(btc_eth_returns[,1]) ~ eth_decile,  # need to convert back to vectors
        main = "% return BTC ~ ETH")


# next step: all altcoins ~ BTC
```

```{r Return_MDay}
# Returns ~ day of the month
# adding day of the month column 
coins_12m$mday <- format(coins_12m$date, "%d")

# calculate BTC returns on original dataframe
btc_returns_vec <- apply(coins_12m[coins_12m$symbol=="BTC (Bitcoin)",c("close")], 2, function (x) { diff(x)/x[-length(x)] }) 
btc_returns_vec <- c(0, btc_returns_vec) # add 0 because first return can't be calculated

# Scatter plot BTC returns ~ day of the month
plot(btc_returns_vec ~ as.matrix(coins_12m[coins_12m$symbol=="BTC (Bitcoin)", "mday"]), 
     xlab = "Day (1-365)", 
     main = "Distribution of BTC returns according day of the month")

# boxplot better
boxplot(btc_returns_vec ~ as.matrix(coins_12m[coins_12m$symbol=="BTC (Bitcoin)", "mday"]), 
     xlab = "Day (1-365)", 
     main = "Distribution of BTC returns according day of the month")
# Conclusion: no strong correlation (idea that returns are higher early in the month not backed by the data)

# is there a better way? (built-in function)
# ggsubseriesplot(as.ts(btc_eth_returns[,1]), # how to show returns per day of the month
#                 main = "BTC ")


```


```{r Log_diff}
# log difference 
btc_eth %>% log %>% diff %>% head

plot(as.zoo(btc_eth %>% log %>% diff), screens = 1, lty = 1:2, xlab = "Date", ylab = "Log Difference")
legend("topleft", c("btc", "eth"), lty = 1:2, cex = 0.5)
```

```{r Candlestick}
# BTC candlestick chart with moving average
candleChart(btc.full, type = "candlesticks", up.col = "black", dn.col = "red", theme = "white")
addSMA(n = c(20, 50, 200))
```

```{r MaxDrawdown}
# PerformanceAnalystics
#library(PerformanceAnalytics)
maxDrawdown(btc_eth_returns)
```

```{r}
# Build Index

# all with numeric vectors
btc_returns <- apply(coins_12m[coins_12m$symbol=="BTC (Bitcoin)",c("close")], 2, function (x) { diff(x)/x[-length(x)] }) 
btc_returns  <- c(0, btc_returns )
 
# mcap_index <- eth_returns * coins_12m[coins_12m$symbol=="ETH",c("market")] +
#   dash_returns * coins_12m[coins_12m$symbol=="DASH",c("market")]

# save 1 df for each coin
TS_list <- list()
for (symb in unique(coins_12m$symbol)) {
  TS_list[[symb]] <- coins_12m[coins_12m$symbol==symb,c("date", "close", "market")]
}
# access one coin's df
# TS_list$BTC

# calculate returns for each df in the TS_list
for (i in 1:length(TS_list)) {
  TS_list[[i]]["close"] <- c(0, apply(TS_list[[i]]["close"], 2, function (x) { ifelse(is.nan(diff(x)/x[-length(x)]) | is.infinite(diff(x)/x[-length(x)]) , 0, diff(x)/x[-length(x)]) }))
}

# library(padr) # to pad with 0 for missing dates, to get elements with same number of rows
# pad(df_name)
# pad(TS_list[[3]], start_val = min(coins_12m$date), end_val = max(coins_12m$date)) # doesn't work
#Reduce("+", lapply(liste, "[[", 2)) # this sums the ith element of each vector of the liste (vectorized)
# to be used if vectors are padded with missing dates (but this is wasteful in terms of memory)

# Initialize weighed index data.frame 
mcap_index <- coins_12m[coins_12m$symbol=="BTC (Bitcoin)", c("date", "close")]
mcap_index$close <- 0
colnames(mcap_index)[2] <- "weighted_return"

# Initialize weighed index data.frame -without BTC
mcap_no_btc_index <- coins_12m[coins_12m$symbol=="BTC (Bitcoin)", c("date", "close")] #initialized w BTC dates
mcap_no_btc_index$close <- 0
colnames(mcap_no_btc_index)[2] <- "weighted_return"

# let's go with the ugly double loop # TO DO: vectorisation. Need regular Time Series (all dates filled)
# works but slow
# for (date_var in days) {
#   accu_wr <- 0 # weighted return
#   accu_w <- 0 # sum of weights
#   for (i in 1:length(TS_list)) {   
#     if (date_var %in% TS_list[[i]]$date) {
#       
#       if (names(TS_list[i])=="BTC (Bitcoin)") { next } # exclude BTC from the index
#       accu_wr = accu_wr + TS_list[[i]][TS_list[[i]]$date==date_var,"close"] * TS_list[[i]][TS_list[[i]]$date==date_var,"market"]
#       accu_w = accu_w + TS_list[[i]][TS_list[[i]]$date==date_var,"market"]
#       if (is.nan(accu_wr$close)) { cat("Non numeric weighted average\n"); cat(i); cat("\n"); break }
#     }
#   }
#   mcap_no_btc_index[mcap_no_btc_index$date==date_var, c("weighted_return")] <- accu_wr/accu_w
# }

# Saving weighted returns df
#save(mcap_index, file = "mcap_index.RData")
# save(mcap_no_btc_index, file = "mcap_no_btc_index.RData")

# Reloading from saved file
# load("./mcap_index.RData")
load("./mcap_no_btc_index.RData")
# mcap_index$weighted_return <- mcap_index$close # fix because wrong column was used
# mcap_index$close = NULL
```

```{r Autocorrelation_BTC, echo = FALSE}
# Autocorrelation (use returns)
ggAcf(TS_list$`BTC (Bitcoin)`[,2], lag.max = 30, type = "correlation")
```

```{r Correlation_index_BTC}
ggCcf(as.vector(mcap_no_btc_index$weighted_return), btc_returns)
```

```{r boxplot_BTC_ETH, echo=FALSE, eval=FALSE}
# btc_decile <- cut(TS_list$`BTC (Bitcoin)`$close, 20)
boxplot(as.vector(mcap_no_btc_index$weighted_return) ~ btc_decile,  # need to convert back to vectors
        main = "% return index ~ BTC")
```

```{r AAPL_test, echo=FALSE, eval=FALSE}
start <- as.Date("2016-01-01")
end <- as.Date("2016-10-01")
getSymbols("AAPL", src = "yahoo", from = start, to = end, adjust = FALSE)

AAPL %>% head
candleChart(AAPL, up.col = "black", dn.col = "red", theme = "white")

```


