---
title: "Analysis of Crypto Currencies"
author: "Quentin Picard"
date: "October 7, 2017"
output: html_document
---

### Introduction

The goal of this document is to provide an overview of crypto currencies, or "coins". The Bitcoin white paper was written in late 2008 but today many more currencies than Bitcoin exist. Some of the questions we will answer include: what are the coins relative importance (by total value)? How have prices behaved in the last few years? How would basic investing strategies have performed during that timeframe? 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r global_options, include=FALSE}
# set up global Knit options for all chunks
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align = "center", include=FALSE, echo=FALSE, warning=FALSE, message=FALSE, cache=T)
```

```{r Libraries}
library(data.table)
library(dplyr)
library(ggplot2)
library(ggvis)
library(ggfortify)
library(xts)
library(quantmod)
library(PerformanceAnalytics)
library(forecast)
```


```{r Data_load}
raw.crypto <- read.csv("./crypto.csv", stringsAsFactors = F) 
#fread("./crypto.csv", stringsAsFactors = F) not working anymore w Knit?
crypto <- raw.crypto 
crypto <- crypto[!is.na(date)] # remove 4 NA records
crypto <- na.omit(crypto)
```
### Data 
The data file studied here contains OHLC (Open-High-Low-Close) data for about 1100 crypto currencies during the last 5 years. The data was scrapped from https://coinmarketcap.com/ by [Jvent](https://www.kaggle.com/jessevent) and uploaded to Kaggle.

Source file: 
[link](https://www.kaggle.com/jessevent/all-crypto-currencies) 

Here is a sample of the data:
```{r Data_Exploration, include=T, echo=F}
crypto %>% head(5)
```

```{r}
# Checking format of imported columns
str(crypto)
```

```{r variable_formats}
# Columns format fix
# crypto[,date:=as.Date(date)]
crypto$date <- as.Date(crypto$date)
crypto$volume <- as.numeric(crypto$volume)
crypto$market <- as.numeric(crypto$market)
```

```{r Duplicates}
# Both coin/ symbol columns have duplicates (e.g. "Bitcoin"/ "ABC")
crypto$symbol <- paste(crypto$symbol, " (", crypto$coin, ")", sep = "")
```

The file contains `r length(levels(as.factor(crypto$symbol)))` different coins. Here are the first 20 in alphabetical order:
```{r Coin_sample, include=T}
# List of coins
l <- levels(as.factor(crypto$symbol)) %>% head(20) 
for (i in l) {
  cat(i)
  cat("\n")
}
```

```{r NAs analysis}
# analysis of NAs
NA.count <- sapply(crypto, function(x) sum(is.na(x)))
NA.percentage <- sapply(crypto, function(x) 100*sum(is.na(x))/length(x))
NA.summary = cbind(NA.count[NA.count>0], NA.percentage[NA.count>0])
NA.summary[order(desc(NA.summary[,2])),] # show by decreasing % of NA's (matrix sorting)
NA.summary[order(-(NA.summary[,2])),] # - means DESC
# Conclusion: all cells have numeric values (but there are zeros!)
```

```{r MissingDates}
# holes in dates
test <- crypto %>% 
  group_by(date) %>%
  summarize() 

test$date <- c(0,diff(test$date)) # date difference between 2 rows
dim(test[test$date != 1,]) # no holes (at least one coin per date)
```

```{r MissingMarketCap}
# 0 market caps despite close price > 0 and some volume
crypto %>%
  filter(market==0) %>% 
  filter(close>0) %>% # price not 0
  filter(volume>1000) %>% # some volume
  dim
# these are probably missing data points
```

```{r SmallMarketCap}
# Calculate figures for commentary just below
null_mc <- dim(crypto[crypto$date=="2017-09-28" & crypto$market==0,])[1]
null_mc
small_mc <- dim(crypto[crypto$date=="2017-09-28" & crypto$market<=1e6,])[1]
small_mc
# Percentages
null_mc_perc <- round(100*dim(crypto[crypto$date=="2017-09-28" & crypto$market==0,])[1]/length(levels(as.factor(crypto$symbol))), 2)
null_mc_perc
small_mc_perc <- round(100*dim(crypto[crypto$date=="2017-09-28" & crypto$market<=1e6,])[1]/length(levels(as.factor(crypto$symbol))), 2)
small_mc_perc
```

### Coins creation
The following graph shows the constant rise of reported coins. We observe some volatility in the number of new coins in last few months. One factor was China's decision to ban initial coin offerings (ICOs) on September, 4th.
Today the pace of new ICOs globally is still strong.
```{r Montly_coin_count, include=T}
# Create data.frame for graph
crypto_active <- crypto %>% 
  group_by(coin) %>% 
  mutate(month=format(date,"%Y%m")) %>%
  mutate(first_month=ifelse(date==min(date),"yes", "no")) %>%
  mutate(last_month=ifelse(date==max(date) & format(date, "%Y%m")!="201709","yes", "no")) %>%
  ungroup() %>%
  group_by(month) %>%
  summarise(active=n_distinct(coin), 
            initiations=sum(first_month=="yes"), 
            discontinuations=sum(last_month=="yes"))

crypto_active %>%
  ggplot(aes(x=month, y=value)) + 
  geom_line(aes(y=active, col="Monthly active coins"), group=1) +
  geom_line(aes(y=initiations, col="Coin initiations"), group=1) +
  coord_cartesian(ylim = c(0, 1100)) +
  ggtitle("Coin initiations (coinmarketcap.com)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_x_discrete(name ="Month", breaks = c("201304","201404","201504", "201604", "201704"), labels = c("Apr 2013","Apr 2014","Apr 2015", "Apr 2016", "Apr 2017")) +
  guides(colour=guide_legend(title="Legend")) +
  # theme(axis.text.x = element_text(angle = 60, hjust = 1)) + # angled text no longer needed
  scale_y_continuous(name ="Number of coins", breaks=seq(0, 1200, by = 200))

```
```{r}
# Consistency check (despite holes in data per next chunck)
sum(crypto_active$initiations-crypto_active$discontinuations) == crypto_active$active[dim(crypto_active)[1]]
```

```{r}
# Analysis: showing some coins are not reported on all months

# unique coins for 2013-11
coins201311 <-crypto %>% 
  mutate(month=format(date,"%Y%m")) %>%
  subset(select=c(12, 1:11)) %>% # put new month column at the beginning
  distinct(month, coin) %>%          # useful to remove extra rows !
  arrange(month, coin) %>%
  filter(month=="201311")

coins201310 <-crypto %>% 
  mutate(month=format(date,"%Y%m")) %>%
  subset(select=c(12, 1:11)) %>% # put new month column at the beginning
  distinct(month, coin) %>% # useful ! 
  arrange(month, coin) %>%
  filter(month=="201310")

difference <- coins201311 %>% anti_join(coins201310, by = "coin")

newcoins201311 <- crypto %>%
  filter(grepl("2013-11", date)) %>%
  distinct(coin)

difference %>% anti_join(newcoins201311, by = "coin") # Bottlecaps not explained by coin initiations

# showing gap months for Bottlecaps (e.g. 2013-10)
crypto %>% 
  mutate(month=format(date,"%Y%m")) %>%
  subset(select=c(12, 1:11)) %>% # put new month column at the beginning
  distinct(month, coin) %>% # useful ! 
  arrange(month, coin) %>%
  filter(coin=="Bottlecaps")
```

### Market Capitalisation Distribution

Looking at the distribution of coin market caps, we observe a large skew towards smaller valuations. 
In fact `r null_mc` (`r null_mc_perc`%) of coins have no market capitalisation as of September 28th, 2017 (the market cap reported by the website is $0). 
`r small_mc` (`r small_mc_perc`%) have a market cap below $1 million.
```{r MarketCap_Hist, include=T, cache=T}
# "Present" Market Cap histogram (latest day)
crypto[crypto$date=="2017-09-28",] %>% 
  ggplot(aes(x = market)) + 
  geom_histogram(binwidth = 1e6) + coord_cartesian(xlim = c(0, 1e8)) +
  ggtitle("Distribution of coin market caps") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_x_continuous(name ="Coin market capitalization", breaks = c(1e6,2.5e7,1e8, 1e9), labels = scales::comma) +
  scale_y_continuous(name ="Number of coins")

# breaks = c(1e6,2.5e7,1e8, 1e9), limits=c("1,000,000","25,000,000","100,000,000", "1,000,000,000"))

# Taking log and removing coins without 0 market cap
# For now commenting log price distribution as unable to explain clearly

# crypto[crypto$date=="2017-09-28" & crypto$market != 0,] %>% 
#   ggplot(aes(x = log(market+1))) + 
#   geom_histogram(binwidth = 0.1) #+ coord_cartesian(xlim = c(55, 70))

```



```{r holdings_cash, cache=T}
# Defining df holdings_cash (used in later analysis)

# setting a volume threshold at which to buy
volume_threshold <- 100000

# min with default trade_date 
Robustmin <- function(x) {if (length(x)>0) min(x) else Sys.Date()+1} # trade date when no trade was made: tomorrow

# define trade date
crypto_trade_date <- crypto %>% 
  group_by(coin) %>%  
  arrange(date, coin) %>%
  mutate(trade_date=Robustmin(date[volume>volume_threshold])) # add trade_date

# get trade price at trade date (close of the day). Check that a date exist and that the price is not 0
crypto_trade_price <- crypto_trade_date %>%
  mutate(trade_price=ifelse(length(close[date==trade_date])==1,ifelse(close[date==trade_date]!=0,close[date==trade_date],Inf),Inf))

# calculate coin holding on any given date
crypto_coin_holding <- crypto_trade_price %>%
  mutate(holding= ifelse(date>=trade_date, close / trade_price, 0)) # holding should be 0 before trade happens

# holdings per coin and total with cash
holdings_cash <- crypto_coin_holding %>%
  group_by(date) %>%
  mutate(cash=-sum(trade_price!=Inf & date >= trade_date)) %>%
  mutate(Total_holdings=sum(holding)+cash) %>%
  mutate(holding_coin_perc=holding/(Total_holdings-cash)*100) %>%  # % of coin assets only (excluding cash)
  arrange(desc(date))
```

### Coins Market Cap graph

This graph shows the evolution of each coin total market cap over time. 
```{r MarketCapGraph, include=T, fig.width=9, fig.height=6}
# Graph - market cap evolution
options("scipen" = 10)

symbol_tooltip <- function(x) {
  if (is.null(x)) return(NULL)
  
  paste0(names(x)[3],": <b>", x[[3]], "</b><br>", # Coin
         "Market Cap: ", format(x[[2]],big.mark=",", trim=TRUE), "<br>", # Market cap
         names(x)[1], ": ", as.Date(x[[1]]/86400000, origin='1970-01-01'), "<br>")
}

# Coin filter
coin_filter <- holdings_cash %>%
  group_by(symbol) %>%
  summarise(MC_threshold=max(market)) %>% 
  filter(MC_threshold>2e9)

#  ************ Coin Market Cap lines with tooltip ************
# Uses both lines and (transparent) points so point give right date
# inner_join(holdings_cash, coin_filter, by= "symbol") %>%
#   ggvis(x = ~date, y = ~market, stroke = ~symbol) %>%
#   ggvis::add_axis("x", title = "Date") %>%
#   ggvis::add_axis("y", title = "Market Cap (Billions)", title_offset = 100) %>%
#   layer_lines(strokeWidth := 1) %>%
#   layer_points(size = 1, fill = ~symbol, opacity := 0) %>%
#   add_tooltip(symbol_tooltip, on = "hover") %>%  
#   hide_legend('fill') %>%
#   hide_legend('stroke')

# Can't seem to publish when this is one (works fine on local browser..)
# reactive({
#   inner_join(holdings_cash, coin_filter, by= "symbol") %>%
#     ggvis(x = ~date, y = ~market, stroke = ~symbol) %>%
#     ggvis::add_axis("x", title = "Date") %>%
#     ggvis::add_axis("y", title = "Market Cap (Billions)", title_offset = 100) %>%
#     layer_lines(strokeWidth := 1) %>%
#     layer_points(size = 1, fill = ~symbol, opacity := 0) %>%
#     add_tooltip(symbol_tooltip, on = "hover") %>%  
#     hide_legend('fill') %>%
#     hide_legend('stroke')
# }) %>% bind_shiny("plot1")
# ggvisOutput("plot1")

df <- inner_join(holdings_cash, coin_filter, by= "symbol")

df %>%
  ggvis(x = ~date, y = ~market, stroke = ~symbol) %>%
  ggvis::add_axis("x", title = "Date") %>%
  ggvis::add_axis("y", title = "Market Cap (Billions)", title_offset = 100) %>%
  layer_lines(strokeWidth := 1) %>%
  layer_points(size = 1, fill = ~symbol, opacity := 0) %>%
  add_tooltip(symbol_tooltip, on = "hover") %>%
  # ggvis::add_legend("Coins") %>%
  # hide_legend('fill') %>%
  hide_legend('stroke')


```

Since a few coins were clearly dominant we will now represent their relative share of the overall crypto currency pie as measured by market cap. This graph shows a key trend in the past year: the emergence of new coins such as Ethereum and Ripple (22% and 5% of total market cap respectively) and a reduction of the Bitcoin total share from 90% to 50%. 
```{r MarketDominance, include=T, message=FALSE}
# Stacked Area Chart

# coin ordering - decreasing Market Cap (last day)
coin_order2 <- inner_join(holdings_cash, coin_filter, by= "symbol") %>% 
  filter(date=="2017-09-28") %>%
  group_by(symbol) %>%
  arrange(date, desc(market)) 

# Plot
inner_join(holdings_cash, coin_filter, by= "symbol") %>% 
  group_by(symbol) %>%
  arrange(date, desc(market)) %>%
  ungroup() %>%
  group_by(date) %>%
  mutate(market_prop = market/sum(market)) %>% 
  ggplot(aes(x=date, y=market_prop, fill=factor(symbol, levels = as.list(coin_order2)$symbol))) + 
  geom_area(alpha=1 , size=1) + 
  ggtitle("Market Cap Dominance") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Date") + 
  scale_y_continuous(name ="Proportion of Total Market Cap") +
  guides(fill=guide_legend(title="Top coins by Market Cap"))

```

### Return distribution

In the graph below, all coins prices were normalized to start trade at $1 on April 28th, 2013.
Plotting prices or log prices for all coins isn't a viable option here. The coins should be clustered so that it's possible to read the data. This analysis is outside the scope of this document.
```{r Graph with normalized start date & price, include=T}

crypto_initiation_date <- crypto %>%  group_by(coin) %>%  mutate(initiation_date=min(date)) # add initiation date

crypto_initiation_date_price <- crypto_initiation_date %>%
  mutate(initiation_price=close[date==initiation_date])  # condition on different line in group ! very useful !

crypto_norm <- crypto_initiation_date_price %>%
  mutate(relative_date=(date-(initiation_date-as.Date("2013-04-28"))), # calculate relative date to normalize starts
         relative_price=ifelse(initiation_price!=0, close/initiation_price, close) ) 
#View(crypto_norm)

crypto_norm_volume <- crypto_norm %>% group_by(coin) %>% mutate(avg_volume=mean(volume)) # average volume per coin
#View(crypto_norm_volume)

library(gridExtra)

crypto_norm_volume[,] %>%
  filter(avg_volume > 100000) %>%
  ggplot() + 
  geom_line(aes(x=relative_date, y=log(relative_price+1), group=symbol)) + 
  labs(x = "Date") + ylab("Coin Prices (Log)") +
  ggtitle("Coin Price Series (Log)") +
  theme(plot.title = element_text(hjust = 0.5))

```

```{r dollar averaging}
# Checks
# Tests on trading date
crypto_trade_date[crypto_trade_date$coin=="808Coin" & 
                    (crypto_trade_date$date == crypto_trade_date$trade_date |
                       crypto_trade_date$date == crypto_trade_date$trade_date -1 |
                       crypto_trade_date$date == crypto_trade_date$trade_date +1),]

# Tests on trading price
crypto_trade_price[crypto_trade_price$symbol=="FLDC" & 
                    (crypto_trade_price$date == crypto_trade_price$trade_date |
                       crypto_trade_price$date == crypto_trade_price$trade_date -1 |
                       crypto_trade_price$date == crypto_trade_price$trade_date +1),]
# Conclusion: would need to look into coins with volume > threshold but price = 0 (e.g. FLDC)
```

```{r Stats}
# % of coins bought
perc <- dim(crypto_trade_date[crypto_trade_date$trade_date!=Sys.Date()+1,])[1]/dim(crypto_trade_date)[1]*100
cat("Percentage of coins bought (based on volume threshold): \n", perc, "\n")
```

```{r Portfolio Pie Chart prep}
# Total Holdings 
holdings <- crypto_coin_holding %>%
  group_by(date) %>%
  filter(date=="2017-09-28") %>%
  mutate(Total_holdings=sum(holding)) %>%
  mutate(holding_coin_perc=holding/Total_holdings*100) %>%
  filter(holding_coin_perc>0.5) %>%
  arrange(desc(holding_coin_perc))

holdings

```

### Trading Strategies
In order to better understand the behavior of these coins, I implemented a basic dollar averaging trading strategy. Starting 5 years ago when data is available, everytime the daily trading volume passes a certain threshold ($100,000 to start) for a given coin, I buy $1 of it and hold it forever. 
The composition of the portfolio on the last day where data is available (September 28th, 2017) is shown below:

```{r DollarAveragingPieChart, include=T, fig.width=9, fig.height=6}
# coin ordering for pie chart
coin_order <- crypto_coin_holding %>%
  group_by(date) %>%
  filter(date=="2017-09-28") %>%
  mutate(Total_holdings=sum(holding)) %>%
  mutate(holding_coin_perc=holding/Total_holdings*100) %>%
  filter(holding_coin_perc>0.5) %>%
  arrange(holding_coin_perc) %>%
  ungroup() %>%
  select(coin)

# need to reorder fill factor levels in inverse order of appearance (see next chunk)
holdings %>% ggplot(aes(x = "", y= holding_coin_perc, fill=factor(coin, levels = as.list(coin_order)$coin))) + 
  geom_bar(stat = "identity") + 
  coord_polar(theta = "y") +
  ggtitle("$ Averaging Final Portfolio") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_discrete(name = "Top coins in Dollar Averaging Portfolio (smallest to largest)") +
  labs(x = "") + ylab("")
```

Conclusion: 50% of the portfolio is made of just 6 coins. Stratis, the biggest position, went up 200,000 times. The next step would be to check if those coins indeed could have been purchased at the initial 'low' prices reported on Coinmarketcap.com. Historical exchange data would be needed for this.

```{r PortfolioReturn}
# Portfolio Return Percentage
return <- -100 *(holdings_cash$Total_holdings[1] - holdings_cash$cash[1]) / holdings_cash$cash[1]
cat("Percentage return: ", return, "\n")
```
The value of the portfolio over time (net of the cash investment represented in red) is plotted below:
```{r PortfolioValue, include=T}
holdings_cash %>%
  filter(holding_coin_perc>1) %>%
  ggplot(aes(x=date)) + 
  geom_line(aes(y=holding, group=symbol)) +
  geom_line(aes(y=Total_holdings, color = "Total Holdings (coins + cash)")) + 
  geom_line(aes(y=cash, color = "Cash invested")) +
  labs(color='Legend') +
  ggtitle("Value of Portfolio over time") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Date") + ylab("Value ($)")
```

```{r Coin holdings ~ volume}
# Function to calculate final portfolio value and return on the invested cash
# TO DO: refactor for space efficiency

Holdings_from_volume <- function(crypto, volume_threshold=100000, symbols_to_exclude=list("")) {
  #returns a list with the total final holding and the percentage return on the invested cash
  
  # define trade date
  crypto_trade_date <- crypto %>% 
    group_by(coin) %>%  
    arrange(date, coin) %>%
    mutate(trade_date=Robustmin(date[volume>volume_threshold])) # add trade_date
  
  # get trade price at trade date (close of the day). Check that a date exist and that the price is not 0
  crypto_trade_price <- crypto_trade_date %>%
    mutate(trade_price=ifelse(length(close[date==trade_date])==1,ifelse(close[date==trade_date]!=0,close[date==trade_date],Inf),Inf))
  
  # calculate coin holding on any given date
  crypto_coin_holding <- crypto_trade_price %>%
    filter(!(symbol %in% symbols_to_exclude)) %>%    # add filter: symbol in arguments
    mutate(holding= ifelse(date>=trade_date, close / trade_price, 0)) # holding should be 0 before trade happens
  
  # holdings per coin and total with cash
  holdings_cash <- crypto_coin_holding %>%
    group_by(date) %>%
    mutate(cash=-sum(trade_price!=Inf & date >= trade_date)) %>%
    mutate(Total_holdings=sum(holding)+cash) %>%
    mutate(holding_coin_perc=holding/(Total_holdings-cash)*100) %>%  # % of coin assets only (excluding cash)
    arrange(desc(date))
  
  # return total holding value (last date)
  # a <- holdings_cash[holdings_cash$date==as.Date("2017-09-28"),c("Total_holdings")] # core R
  total_holding <- holdings_cash %>%  distinct(date,Total_holdings)
  
  #cat("Total holdings: \n", total_holding$Total_holdings[1], "\n")
  c(as.numeric(total_holding$Total_holdings[1]),
    -100 *(holdings_cash$Total_holdings[1] - holdings_cash$cash[1]) / holdings_cash$cash[1])
}

Holdings_from_volume(crypto, 100000, "")
Holdings_from_volume(crypto, 100000, list("BTC", "ETH"))


```

Finally I plotted the value of the final portfolio as a function of the volume threshold at which I would buy each coin.
The gains get exponentially higher as the threshold diminishes. Meaning that if I buy each coin "right away" -say when the daily volume is only above $1000- then some of the coins will experience a meteoric rise and the portfolio will grow more during the holding period.
Same conclusion here: we would need to check with exchange data if these coins were truly available at such small volumes. At the very least we can conclude that this trading strategy would only be available to small players.

```{r VolumeGraph, include=T, cache=T}
# Total holdings obtained as a function of volume treshold
n <- 0:9
vol_vec <- c(1000*((3)^n))
hold_vec <- sapply(vol_vec, function(x) {Holdings_from_volume(crypto, x)}) # calculate final holdings

df <- data.frame(vol_vec,hold_vec[1,1:length(vol_vec)],hold_vec[2,1:length(vol_vec)])
ggplot(df, aes(vol_vec)) + 
  geom_line(aes(y=hold_vec[2,1:length(vol_vec)]), colour="black") + 
  ggtitle("Portfolio Return Relative to Minimum Buy Volume") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Total Return (%)") + 
  xlab("Buy Volume Threshold")
```

```{r self_contained = FALSE}
# Time Series

# filter to last 12 months of data
coins_12m <- holdings_cash %>%
  filter(date > (as.Date("2017-09-28")-365)) %>%
  arrange(date)

# zoo/ xts
days <- seq(min(coins_12m$date), max(coins_12m$date), by = "day")
btc <- zoo(coins_12m[coins_12m$symbol=="BTC (Bitcoin)", "close"], days)
btc.full <- zoo(coins_12m[coins_12m$symbol=="BTC (Bitcoin)", c("open", "high", "low", "close", "volume")], days) # with all info
eth <- zoo(coins_12m[coins_12m$symbol=="ETH (Ethereum)", "close"], days)

autoplot(btc, main = "Bitcoin price - Past year")
plot(btc, main = "Bitcoin price - Past year")

btc_eth <- as.xts(data.frame(btc_col = btc, 
                             eth_col = eth))

colnames(btc_eth) <- c("btc_close", "eth_close")
head(btc_eth)

plot(as.zoo(btc_eth), screens = 1, lty = 1:2, xlab = "Date", ylab = "Closing price")
legend("topleft", c("btc", "eth"), lty = 1:2, cex = 0.5)

# ts - plot works but x axis to fix
# btc_ts <- ts(coins_12m[coins_12m$symbol=="BTC", "close"],
#            start = c(2016, as.numeric(format(days[1], "%j"))),
#            frequency = 365)
# autoplot(btc_ts) # x axis format to fix

```

```{r Returns}
# calculate returns - Method #1
quantmod::Delt(btc_eth[, "btc_close"]) %>% head #  period pct. price change - is xts
quantmod::Delt(as.xts(btc[, "close"])) %>% head #  period pct. price change

# method #2 - vectorized formula
n <- nrow(btc)
apply(btc, 2, function (x) { diff(x)/x[-length(x)] }) %>% head

# For n coins (method #2)
btc_eth_returns <- as.xts(apply(btc_eth, 2, function (x) { diff(x)/x[-length(x)] }))
btc_eth_returns %>% head

plot(as.zoo(btc_eth_returns), screens = 1, lty = 1:2, xlab = "Date", ylab = "Return")
legend("topleft", c("Bitcoin", "Ether"), lty = 1:2, cex = 0.5)


```

```{r}
# boxplot BTC ~ ETH 
eth_decile <- cut(as.vector(btc_eth_returns[,2]), 10)
boxplot(as.vector(btc_eth_returns[,1]) ~ eth_decile,  # need to convert back to vectors
        main = "% return BTC ~ ETH")


# next step: all altcoins ~ BTC
```

### Additional studies

In response to comments from the author of the data file, I created 2 more graphs. The first one looked into the alleged strong returns of Bitcoin in the first days of the month. This is not supported by the data which shows no relationship.
Note that I focused on the last 12 months for this analysis as there was limited activity in the previous 4 years.

```{r Return_MDay, include=T}
# Returns ~ day of the month
# adding day of the month column 
coins_12m$mday <- format(coins_12m$date, "%d")

# calculate BTC returns on original dataframe
btc_returns_vec <- apply(coins_12m[coins_12m$symbol=="BTC (Bitcoin)",c("close")], 2, function (x) { diff(x)/x[-length(x)] }) 
btc_returns_vec <- c(0, btc_returns_vec) # add 0 because first return can't be calculated

# Scatter plot BTC returns ~ day of the month
# plot(btc_returns_vec ~ as.matrix(coins_12m[coins_12m$symbol=="BTC (Bitcoin)", "mday"]), 
#      xlab = "Day (1-365)", 
#      main = "Distribution of BTC returns according day of the month")

# boxplot better
boxplot(btc_returns_vec ~ as.matrix(coins_12m[coins_12m$symbol=="BTC (Bitcoin)", "mday"]), 
     xlab = "Day of the month", 
     ylab = "Daily return",
     main = "Distribution of BTC returns according day of the month")
# Conclusion: no strong correlation (idea that returns are higher early in the month not backed by the data)

# is there a better way? (built-in function)
# ggsubseriesplot(as.ts(btc_eth_returns[,1]), # how to show returns per day of the month
#                 main = "BTC ")


```


```{r Log_diff}
# log difference 
btc_eth %>% log %>% diff %>% head

plot(as.zoo(btc_eth %>% log %>% diff), screens = 1, lty = 1:2, xlab = "Date", ylab = "Log Difference")
legend("topleft", c("btc", "eth"), lty = 1:2, cex = 0.5)
```

```{r Candlestick}
# BTC candlestick chart with moving average
candleChart(btc.full, type = "candlesticks", up.col = "black", dn.col = "red", theme = "white")
addSMA(n = c(20, 50, 200))
```

```{r MaxDrawdown}
# PerformanceAnalystics
#library(PerformanceAnalytics)
maxDrawdown(btc_eth_returns)
```

```{r}
# Build Market Cap Weighted Index

# all with numeric vectors
btc_returns <- apply(coins_12m[coins_12m$symbol=="BTC (Bitcoin)",c("close")], 2, function (x) { diff(x)/x[-length(x)] }) 
btc_returns  <- c(0, btc_returns )
 
# mcap_index <- eth_returns * coins_12m[coins_12m$symbol=="ETH",c("market")] +
#   dash_returns * coins_12m[coins_12m$symbol=="DASH",c("market")]

# save 1 df for each coin
TS_list <- list()
for (symb in unique(coins_12m$symbol)) {
  TS_list[[symb]] <- coins_12m[coins_12m$symbol==symb,c("date", "close", "market")]
}
# access one coin's df
# TS_list$BTC

# calculate returns for each df in the TS_list
for (i in 1:length(TS_list)) {
  TS_list[[i]]["close"] <- c(0, apply(TS_list[[i]]["close"], 2, function (x) { ifelse(is.nan(diff(x)/x[-length(x)]) | is.infinite(diff(x)/x[-length(x)]) , 0, diff(x)/x[-length(x)]) }))
}

# library(padr) # to pad with 0 for missing dates, to get elements with same number of rows
# pad(df_name)
# pad(TS_list[[3]], start_val = min(coins_12m$date), end_val = max(coins_12m$date)) # doesn't work
#Reduce("+", lapply(liste, "[[", 2)) # this sums the ith element of each vector of the liste (vectorized)
# to be used if vectors are padded with missing dates (but this is wasteful in terms of memory)

# Initialize weighed index data.frame 
mcap_index <- coins_12m[coins_12m$symbol=="BTC (Bitcoin)", c("date", "close")]
mcap_index$close <- 0
colnames(mcap_index)[2] <- "weighted_return"

# Initialize weighed index data.frame -without BTC
mcap_no_btc_index <- coins_12m[coins_12m$symbol=="BTC (Bitcoin)", c("date", "close")] #initialized w BTC dates
mcap_no_btc_index$close <- 0
colnames(mcap_no_btc_index)[2] <- "weighted_return"

# let's go with the ugly double loop (works but slow)
# TO DO: vectorisation. Need regular Time Series (all dates filled)

# for (date_var in days) {
#   accu_wr <- 0 # weighted return
#   accu_w <- 0 # sum of weights
#   for (i in 1:length(TS_list)) {   
#     if (date_var %in% TS_list[[i]]$date) {
#       
#       if (names(TS_list[i])=="BTC (Bitcoin)") { next } # exclude BTC from the index
#       accu_wr = accu_wr + TS_list[[i]][TS_list[[i]]$date==date_var,"close"] * TS_list[[i]][TS_list[[i]]$date==date_var,"market"]
#       accu_w = accu_w + TS_list[[i]][TS_list[[i]]$date==date_var,"market"]
#       if (is.nan(accu_wr$close)) { cat("Non numeric weighted average\n"); cat(i); cat("\n"); break }
#     }
#   }
#   mcap_no_btc_index[mcap_no_btc_index$date==date_var, c("weighted_return")] <- accu_wr/accu_w
# }

# Saving weighted returns df
#save(mcap_index, file = "mcap_index.RData")
# save(mcap_no_btc_index, file = "mcap_no_btc_index.RData")

# Reloading from saved file
# load("./mcap_index.RData")
load("./mcap_no_btc_index.RData")
# mcap_index$weighted_return <- mcap_index$close # fix because wrong column was used
# mcap_index$close = NULL
```

```{r Autocorrelation_BTC, echo = FALSE}
# Autocorrelation (use returns)
ggAcf(TS_list$`BTC (Bitcoin)`[,2], lag.max = 30, type = "correlation")
```

```{r Correlation_index_BTC}
ggCcf(as.vector(mcap_no_btc_index$weighted_return), btc_returns)
# Conclusion: correction between Bitcoin and rest of the index is clearly visible (no lag)
```

The second shows the returns of a coins index (with 'traditional' market cap weights) in relation to Bitcoin returns. Supposedly Bitcoin price surges were causing alternative coin holders to sell their holdings in exchange for Bitcoin. We see that in fact all crypto currencies seem to move in tandem.
```{r boxplot_index_BTC, include=T}
btc_decile <- cut(TS_list$`BTC (Bitcoin)`$close, 20)
boxplot(as.vector(mcap_no_btc_index$weighted_return) ~ btc_decile,  # need to convert back to vectors
        main = "% Return Weighted Index ~ Bitcoin",
        xlab = "Bitcoin Daily Return Range", 
        ylab = "Weighted Index Daily Return")
# axis(side = 1, at=c(1,5, 10, 15, 20), labels=c("-18.8%", "-10.2%", "0%", "11.1%", "21.8%"))
```

```{r AAPL_test, echo=FALSE, eval=FALSE}
start <- as.Date("2016-01-01")
end <- as.Date("2016-10-01")
getSymbols("AAPL", src = "yahoo", from = start, to = end, adjust = FALSE)

AAPL %>% head
candleChart(AAPL, up.col = "black", dn.col = "red", theme = "white")

```


